
%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\def\COM#1{{\color{red}[#1]}}


\input{epig-macros}

\def\modCG#1{\textcolor{red}{#1}}


%pour le texte dans le tableau

\newcolumntype{M}[1]{>{\raggedright}m{#1}}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
\urldef{\mailsc}\path|louis.fippo-fitime@irccyn.ec-nantes.fr|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}


\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Integrating time-series data in large-scale discrete cell-based models}

% a short form should be given in case it is too long for the running head
\titlerunning{Integrating time-series data in cell-based discrete models}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Louis Fippo Fitime\inst{1}%
\and Christian Schuster\inst{2} \and Peter Angel\inst{2}
\and Olivier Roux\inst{1}\and Carito Guziolowski\inst{1}}
%
\authorrunning{Louis Fippo Fitime et al.}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published

\institute{LUNAM Universit\'e, \'Ecole Centrale de Nantes, IRCCyN UMR CNRS 6597\\
(Institut de Recherche en Communications et Cybern\'etique de Nantes)\\
1 rue de la No\"e -- B.P. 92101 -- 44321 Nantes Cedex 3, France.\\
\mailsc\\
\url{http://www.irccyn.ec-nantes.fr/en/}
\and
Division of Signal Transduction and Growth Control (A100), DKFZ-ZMBH Alliance, Deutsches Krebsforschungszentrum, Heidelberg, Germany
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle


\begin{abstract}
In this work we propose an automatic way of generating and verifying 
formal hybrid models of signaling and transcriptional
events, gathered in large-scale regulatory networks.This is done by 
integrating temporal and stochastic aspects of
the expression of some biological components. The hybrid approach lies 
in the fact that measurements take into account both times of 
lengthening phases and discrete switches between them. The model 
proposed is based on a real case study of keratinocytes differentiation, 
in which gene time-series data was generated upon Calcium stimulation.

To achieve this we rely on the Process Hitting (PH) formalism that was 
designed to consider large-scale system analysis.  We first propose an 
automatic way of detecting and translating biological motifs from  the 
Pathway Interaction Database to the PH formalism. Then, we propose a way of estimating 
temporal and stochastic parameters from time-series expression data of 
action on the PH. Simulations emphasize the interest of synchronizing 
concurrent events.
\keywords{time-series data, large-scale network, hybrid models, compositional approach, stochastic simulation}
\end{abstract}


\input{partsModif/intro}
\input{partsModif/methods}
\input{partsModif/results}
\input{partsModif/conclusion}
\input{partsModif/acknowledgements}
\input{partsModif/Answer_questions}


%\bibliographystyle{plain}
%\bibliography{biblio.bib}

\begin{thebibliography}{4}

\bibitem{batt2005validation}
Gr{\'e}gory Batt, Delphine Ropers, Hidde De~Jong, Johannes Geiselmann, Radu
  Mateescu, Michel Page, and Dominique Schneider.
\newblock Validation of qualitative models of genetic regulatory networks by
  model checking: Analysis of the nutritional stress response in escherichia
  coli.
\newblock {\em Bioinformatics}, 21(suppl 1):i19--i28, 2005.

\bibitem{guziolowski2012automatic}
Carito Guziolowski, Aristotelis Kittas, Florian Dittmann, and Niels Grabe.
\newblock Automatic generation of causal networks linking growth factor stimuli
  to functional cell state changes.
\newblock {\em FEBS Journal}, 279(18):3462--3474, 2012.

\bibitem{guziolowski2013exhaustively}
Carito Guziolowski, Santiago Videla, Federica Eduati, Sven Thiele, Thomas
  Cokelaer, Anne Siegel, and Julio Saez-Rodriguez.
\newblock Exhaustively characterizing feasible logic models of a signaling
  network using answer set programming.
\newblock {\em Bioinformatics}, 29(18):2320--2326, 2013.

\bibitem{heiner2008petri}
Monika Heiner, David Gilbert, and Robin Donaldson.
\newblock Petri nets for systems and synthetic biology.
\newblock In {\em Formal methods for computational systems biology}, pages
  215--264. Springer, 2008.

\bibitem{Kolly2005}
C.~Kolly, M.~M. Suter, and E.~J. Muller.
\newblock {{P}roliferation, cell cycle exit, and onset of terminal
  differentiation in cultured keratinocytes: pre-programmed pathways in control
  of {C}-{M}yc and {N}otch1 prevail over extracellular calcium signals}.
\newblock {\em J. Invest. Dermatol.}, 124(5):1014--1025, May 2005.

\bibitem{macnamara2012state}
Aidan MacNamara, Camille Terfve, David Henriques, Beatriz~Pe{\~n}alver
  Bernab{\'e}, and Julio Saez-Rodriguez.
\newblock State--time spectrum of signal transduction logic models.
\newblock {\em Physical Biology}, 9(4):045003, 2012.

\bibitem{maurin2009modeling}
Myl{\`e}ne Maurin, Morgan Magnin, and Olivier Roux.
\newblock Modeling of genetic regulatory network in stochastic $\pi$-calculus.
\newblock In {\em Bioinformatics and Computational Biology}, pages 282--294.
  Springer, 2009.

\bibitem{mitsos2009identifying}
Alexander Mitsos, Ioannis~N Melas, Paraskeuas Siminelakis, Aikaterini~D
  Chairakaki, Julio Saez-Rodriguez, and Leonidas~G Alexopoulos.
\newblock Identifying drug effects via pathway alterations using an integer
  linear programming optimization formulation on phosphoproteomic data.
\newblock {\em PLoS computational biology}, 5(12):e1000591, 2009.

\bibitem{mobashir2012simulated}
Mohammad Mobashir, Burkhart Schraven, and Tilo Beyer.
\newblock Simulated evolution of signal transduction networks.
\newblock {\em PloS one}, 7(12):e50905, 2012.

\bibitem{molloy1982performance}
Michael~K. Molloy.
\newblock Performance analysis using stochastic petri nets.
\newblock {\em Computers, IEEE Transactions on}, 100(9):913--917, 1982.

\bibitem{PMR10-TCSB}
{L}o{\"i}c {P}aulev{\'e}, {M}organ {M}agnin, and {O}livier {R}oux.
\newblock Refining dynamics of gene regulatory networks in a stochastic
  $\pi$-calculus framework.
\newblock In {\em Transactions on Computational Systems Biology XIII}, pages
  171--191. Springer, 2011.

\bibitem{priami1995stochastic}
Corrado Priami.
\newblock Stochastic $\pi$-calculus.
\newblock {\em The Computer Journal}, 38(7):578--589, 1995.

\bibitem{schaefer2009pid}
Carl~F Schaefer, Kira Anthony, Shiva Krupa, Jeffrey Buchoff, Matthew Day, Timo
  Hannay, and Kenneth~H Buetow.
\newblock Pid: the pathway interaction database.
\newblock {\em Nucleic acids research}, 37(suppl 1):D674--D679, 2009.

\bibitem{Tu2011}
C.~L. Tu, W.~Chang, and D.~D. Bikle.
\newblock {{T}he calcium-sensing receptor-dependent regulation of cell-cell
  adhesion and keratinocyte differentiation requires {R}ho and filamin {A}}.
\newblock {\em J. Invest. Dermatol.}, 131(5):1119--1128, May 2011.

\bibitem{tyson2003sniffers}
John~J Tyson, Katherine~C Chen, and Bela Novak.
\newblock Sniffers, buzzers, toggles and blinkers: dynamics of regulatory and
  signaling pathways in the cell.
\newblock {\em Current opinion in cell biology}, 15(2):221--231, 2003.

\bibitem{yu2012hipathdb}
Namhee Yu, Jihae Seo, Kyoohyoung Rho, Yeongjun Jang, Jinah Park, Wan~Kyu Kim,
  and Sanghyuk Lee.
\newblock hipathdb: a human-integrated pathway database with facile
  visualization.
\newblock {\em Nucleic acids research}, 40(D1):D797--D802, 2012.

\end{thebibliography}


\end{document}
