
%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\input{epig-macros}

\def\modCG#1{\textcolor{red}{#1}}


%pour le texte dans le tableau

\newcolumntype{M}[1]{>{\raggedright}m{#1}}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
\urldef{\mailsc}\path|louis.fippo-fitime}@irccyn.ec-nantes.fr|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}


\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Integrating time-series data in large-scale discrete cell-based models}

% a short form should be given in case it is too long for the running head
\titlerunning{Integrating time-series data in cell-based discrete models}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Louis Fippo Fitime\inst{1}%
\and Christian Schuster\inst{2} \and Peter Angel\inst{2}
\and Olivier Roux\inst{1}\and Carito Guziolowski\inst{1}}
%
\authorrunning{Louis Fippo Fitime et al.}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published

\institute{LUNAM Universit\'e, \'Ecole Centrale de Nantes, IRCCyN UMR CNRS 6597\\
(Institut de Recherche en Communications et Cybern\'etique de Nantes)\\
1 rue de la No\"e -- B.P. 92101 -- 44321 Nantes Cedex 3, France.\\
\mailsc\\
\url{http://www.irccyn.ec-nantes.fr/en/}
\and
Division of Signal Transduction and Growth Control (A100), DKFZ-ZMBH Alliance, Deutsches Krebsforschungszentrum, Heidelberg, Germany
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle


\begin{abstract}
In this work we propose an automatic way of generating and verifying 
formal hybrid models of signaling and transcriptional
events, gathered in large-scale regulatory networks.This is done by 
integrating temporal and stochastic aspects of
the expression of some biological components. The hybrid approach lies 
in the fact that measurements take into account both times of 
lengthening phases and discrete switches between them. The model 
proposed is based on a real case study of keratinocytes differentiation, 
in which gene time-series data was generated upon Calcium stimulation.

To achieve this we rely on the Process Hitting (PH) formalism that was 
designed to consider large-scale system analysis.  We first propose an 
automatic way of detecting and translating biological motifs from  the 
PID database to the PH formalism. Then, we propose a way of estimating 
temporal and stochastic parameters from time-series expression data of 
action on the PH. Simulations emphasize the interest of synchronizing 
concurrent events.
\keywords{time-series data, large-scale network, hybrid models, compositional approach, stochastic simulation}
\end{abstract}


\input{parts/intro}
\input{parts/methods}
\input{parts/results}
\input{parts/conclusion}
\input{parts/acknowledgements}



\begin{thebibliography}{4}

\bibitem{schaefer2009pid}
C.~F. Schaefer, K.~Anthony, S.~Krupa, J.~Buchoff, M.~Day, T.~Hannay, and K.~H.
  Buetow, ``Pid: the pathway interaction database,'' \emph{Nucleic acids
  research}, vol.~37, no. suppl 1, pp. D674--D679, 2009.

\bibitem{yu2012hipathdb}
N.~Yu, J.~Seo, K.~Rho, Y.~Jang, J.~Park, W.~K. Kim, and S.~Lee, ``hipathdb: a
  human-integrated pathway database with facile visualization,'' \emph{Nucleic
  acids research}, vol.~40, no.~D1, pp. D797--D802, 2012.

\bibitem{tyson2003sniffers}
J.~J. Tyson, K.~C. Chen, and B.~Novak, ``Sniffers, buzzers, toggles and
  blinkers: dynamics of regulatory and signaling pathways in the cell,''
  \emph{Current opinion in cell biology}, vol.~15, no.~2, pp. 221--231, 2003.

\bibitem{batt2005validation}
G.~Batt, D.~Ropers, H.~De~Jong, J.~Geiselmann, R.~Mateescu, M.~Page, and
  D.~Schneider, ``Validation of qualitative models of genetic regulatory
  networks by model checking: Analysis of the nutritional stress response in
  escherichia coli,'' \emph{Bioinformatics}, vol.~21, no. suppl 1, pp.
  i19--i28, 2005.

\bibitem{mobashir2012simulated}
M.~Mobashir, B.~Schraven, and T.~Beyer, ``Simulated evolution of signal
  transduction networks,'' \emph{PloS one}, vol.~7, no.~12, p. e50905, 2012.

\bibitem{guziolowski2013exhaustively}
C.~Guziolowski, S.~Videla, F.~Eduati, S.~Thiele, T.~Cokelaer, A.~Siegel, and
  J.~Saez-Rodriguez, ``Exhaustively characterizing feasible logic models of a
  signaling network using answer set programming,'' \emph{Bioinformatics},
  vol.~29, no.~18, pp. 2320--2326, 2013.

\bibitem{mitsos2009identifying}
A.~Mitsos, I.~N. Melas, P.~Siminelakis, A.~D. Chairakaki, J.~Saez-Rodriguez,
  and L.~G. Alexopoulos, ``Identifying drug effects via pathway alterations
  using an integer linear programming optimization formulation on
  phosphoproteomic data,'' \emph{PLoS computational biology}, vol.~5, no.~12,
  p. e1000591, 2009.

\bibitem{macnamara2012state}
A.~MacNamara, C.~Terfve, D.~Henriques, B.~P. Bernab{\'e}, and
  J.~Saez-Rodriguez, ``State--time spectrum of signal transduction logic
  models,'' \emph{Physical Biology}, vol.~9, no.~4, p. 045003, 2012.

\bibitem{molloy1982performance}
M.~K. Molloy, ``Performance analysis using stochastic petri nets,''
  \emph{Computers, IEEE Transactions on}, vol. 100, no.~9, pp. 913--917, 1982.

\bibitem{heiner2008petri}
M.~Heiner, D.~Gilbert, and R.~Donaldson, ``Petri nets for systems and synthetic
  biology,'' in \emph{Formal methods for computational systems biology}.\hskip
  1em plus 0.5em minus 0.4em\relax Springer, 2008, pp. 215--264.

\bibitem{priami1995stochastic}
C.~Priami, ``Stochastic $\pi$-calculus,'' \emph{The Computer Journal}, vol.~38,
  no.~7, pp. 578--589, 1995.

\bibitem{maurin2009modeling}
M.~Maurin, M.~Magnin, and O.~Roux, ``Modeling of genetic regulatory network in
  stochastic $\pi$-calculus,'' in \emph{Bioinformatics and Computational
  Biology}.\hskip 1em plus 0.5em minus 0.4em\relax Springer, 2009, pp.
  282--294.

\bibitem{PMR10-TCSB}
L.~{P}aulev{\'e}, M.~{M}agnin, and O.~{R}oux, ``Refining dynamics of gene
  regulatory networks in a stochastic $\pi$-calculus framework,'' in
  \emph{Transactions on Computational Systems Biology XIII}.\hskip 1em plus
  0.5em minus 0.4em\relax Springer, 2011, pp. 171--191.

\bibitem{guziolowski2012automatic}
C.~Guziolowski, A.~Kittas, F.~Dittmann, and N.~Grabe, ``Automatic generation of
  causal networks linking growth factor stimuli to functional cell state
  changes,'' \emph{FEBS Journal}, vol. 279, no.~18, pp. 3462--3474, 2012.


\end{thebibliography}

\end{document}
