
%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{float}

\def\COM#1{{\color{red}[#1]}}

\usepackage{array}  %pour l'allignement dans les tableaux 


\input{epig-macros}
\def\layersep{1cm}

\def\modCG#1{\textcolor{red}{#1}}


%pour le texte dans le tableau

\newcolumntype{M}[1]{>{\raggedright}m{#1}}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
\urldef{\mailsc}\path|louis.fippo-fitime@irccyn.ec-nantes.fr|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}


\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Integrating time-series data in large-scale discrete cell-based models}

% a short form should be given in case it is too long for the running head
\titlerunning{Integrating time-series data in cell-based discrete models}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Louis Fippo Fitime\inst{1}%
\and Christian Schuster\inst{2} \and Peter Angel\inst{2}
\and Olivier Roux\inst{1}\and Carito Guziolowski\inst{1}}
%
\authorrunning{Louis Fippo Fitime et al.}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published

\institute{LUNAM Universit\'e, \'Ecole Centrale de Nantes, IRCCyN UMR CNRS 6597\\
(Institut de Recherche en Communications et Cybern\'etique de Nantes)\\
1 rue de la No\"e -- B.P. 92101 -- 44321 Nantes Cedex 3, France.\\
\mailsc\\
\url{http://www.irccyn.ec-nantes.fr/en/}
\and
Division of Signal Transduction and Growth Control (A100), DKFZ-ZMBH Alliance, Deutsches Krebsforschungszentrum, Heidelberg, Germany
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle


\begin{abstract}
In this work we propose an automatic way of generating and verifying 
formal hybrid models of signaling and transcriptional
events, gathered in large-scale regulatory networks.This is done by 
integrating temporal and stochastic aspects of
the expression of some biological components. The hybrid approach lies 
in the fact that measurements take into account both times of 
lengthening phases and discrete switches between them. The model 
proposed is based on a real case study of keratinocytes differentiation, 
in which gene time-series data was generated upon Calcium stimulation.

To achieve this we rely on the Process Hitting (PH) formalism that was 
designed to consider large-scale system analysis.  We first propose an 
automatic way of detecting and translating biological motifs from  the 
Pathway Interaction Database to the PH formalism. Then, we propose a way of estimating 
temporal and stochastic parameters from time-series expression data of 
action on the PH. Simulations emphasize the interest of synchronizing 
concurrent events.
\keywords{time-series data, large-scale network, hybrid models, compositional approach, stochastic simulation}
\end{abstract}


\input{partsFinal1/intro}
\input{partsFinal1/methods}
\input{partsFinal1/results}
\input{partsFinal1/conclusion}
\input{partsFinal1/acknowledgements}

%\bibliographystyle{plain}
%\bibliography{biblio.bib}
\begin{thebibliography}{10}

\bibitem{Ahmad08}
Jamil Ahmad, Olivier Roux, Gilles Bernot, Jean-Paul Comet, and Adrien Richard.
\newblock Analysing formal models of genetic regulatory networks with delays.
\newblock {\em International Journal of Bioinformatics Research and
  Applications (IJBRA)}, 4(3):240--262, 2008.

\bibitem{AMSW2012}
David~Spieler Aleksandr~Andreychenko, Linar~Mikeev and Verena Wolf.
\newblock Approximate maximum likelihood estimation for stochastic chemical
  kinetics.
\newblock {\em EURASIP Journal on Bioinformatics and Systems Biology}.

\bibitem{Batt15092010}
Gregory Batt, Michel Page, Irene Cantone, Gregor Goessler, Pedro Monteiro, and
  Hidde de~Jong.
\newblock Efficient parameter search for qualitative models of regulatory
  networks using symbolic model checking.
\newblock {\em Bioinformatics}, 26(18):i603--i610, 2010.

\bibitem{batt2007timed}
Gr{\'e}gory Batt, Ramzi~Ben Salah, and Oded Maler.
\newblock {\em On timed models of gene networks}.
\newblock Springer, 2007.

\bibitem{busch2008gene}
Hauke Busch, David Camacho-Trullio, Zbigniew Rogon, Kai Breuhahn, Peter Angel,
  Roland Eils, and Axel Szabowski.
\newblock Gene network dynamics controlling keratinocyte migration.
\newblock {\em Molecular systems biology}, 4(1):199, 2008.

\bibitem{chaouiya2003qualitative}
Claudine Chaouiya, Elisabeth Remy, Brigitte Moss{\'e}, and Denis Thieffry.
\newblock Qualitative analysis of regulatory graphs: a computational tool based
  on a discrete formal framework.
\newblock In {\em Positive Systems}, pages 119--126. Springer, 2003.

\bibitem{de2003genetic}
Hidde De~Jong, Johannes Geiselmann, C{\'e}line Hernandez, and Michel Page.
\newblock Genetic network analyzer: qualitative simulation of genetic
  regulatory networks.
\newblock {\em Bioinformatics}, 19(3):336--344, 2003.

\bibitem{gardner2003inferring}
Timothy~S Gardner, Diego Di~Bernardo, David Lorenz, and James~J Collins.
\newblock Inferring genetic networks and identifying compound mode of action
  via expression profiling.
\newblock {\em Science}, 301(5629):102--105, 2003.

\bibitem{guziolowski2012automatic}
Carito Guziolowski, Aristotelis Kittas, Florian Dittmann, and Niels Grabe.
\newblock Automatic generation of causal networks linking growth factor stimuli
  to functional cell state changes.
\newblock {\em FEBS Journal}, 279(18):3462--3474, 2012.

\bibitem{guziolowski2013exhaustively}
Carito Guziolowski, Santiago Videla, Federica Eduati, Sven Thiele, Thomas
  Cokelaer, Anne Siegel, and Julio Saez-Rodriguez.
\newblock Exhaustively characterizing feasible logic models of a signaling
  network using answer set programming.
\newblock {\em Bioinformatics}, 29(18):2320--2326, 2013.

\bibitem{heiner2008petri}
Monika Heiner, David Gilbert, and Robin Donaldson.
\newblock Petri nets for systems and synthetic biology.
\newblock In {\em Formal methods for computational systems biology}, pages
  215--264. Springer, 2008.

\bibitem{kauffman1969metabolic}
Stuart~A Kauffman.
\newblock Metabolic stability and epigenesis in randomly constructed genetic
  nets.
\newblock {\em Journal of theoretical biology}, 22(3):437--467, 1969.

\bibitem{Kolly2005}
C.~Kolly, M.~M. Suter, and E.~J. Muller.
\newblock {{P}roliferation, cell cycle exit, and onset of terminal
  differentiation in cultured keratinocytes: pre-programmed pathways in control
  of {C}-{M}yc and {N}otch1 prevail over extracellular calcium signals}.
\newblock {\em J. Invest. Dermatol.}, 124(5):1014--1025, May 2005.

\bibitem{macnamara2012state}
Aidan MacNamara, Camille Terfve, David Henriques, Beatriz~Pe{\~n}alver
  Bernab{\'e}, and Julio Saez-Rodriguez.
\newblock State--time spectrum of signal transduction logic models.
\newblock {\em Physical Biology}, 9(4):045003, 2012.

\bibitem{maurin2009modeling}
Myl{\`e}ne Maurin, Morgan Magnin, and Olivier Roux.
\newblock Modeling of genetic regulatory network in stochastic $\pi$-calculus.
\newblock In {\em Bioinformatics and Computational Biology}, pages 282--294.
  Springer, 2009.

\bibitem{mitsos2009identifying}
Alexander Mitsos, Ioannis~N Melas, Paraskeuas Siminelakis, Aikaterini~D
  Chairakaki, Julio Saez-Rodriguez, and Leonidas~G Alexopoulos.
\newblock Identifying drug effects via pathway alterations using an integer
  linear programming optimization formulation on phosphoproteomic data.
\newblock {\em PLoS computational biology}, 5(12):e1000591, 2009.

\bibitem{molloy1982performance}
Michael~K. Molloy.
\newblock Performance analysis using stochastic petri nets.
\newblock {\em Computers, IEEE Transactions on}, 100(9):913--917, 1982.

\bibitem{PMR10-TCSB}
{L}o{\"i}c {P}aulev{\'e}, {M}organ {M}agnin, and {O}livier {R}oux.
\newblock Refining dynamics of gene regulatory networks in a stochastic
  $\pi$-calculus framework.
\newblock In {\em Transactions on Computational Systems Biology XIII}, pages
  171--191. Springer, 2011.

\bibitem{Pinna2010}
Andrea Pinna, Nicola Soranzo, and Alberto de~la Fuente.
\newblock From knockouts to networks: Establishing direct cause-effect
  relationships through graph analysis.
\newblock {\em PLoS ONE}, 5(10):e12912, 10 2010.

\bibitem{porreca2010identification}
Riccardo Porreca, Eugenio Cinquemani, John Lygeros, and Giancarlo
  Ferrari-Trecate.
\newblock Identification of genetic network dynamics with unate structure.
\newblock {\em Bioinformatics}, 26(9):1239--1245, 2010.

\bibitem{priami1995stochastic}
Corrado Priami.
\newblock Stochastic $\pi$-calculus.
\newblock {\em The Computer Journal}, 38(7):578--589, 1995.

\bibitem{RAT2006}
R.~Altman S.~Reinker and J.~Timmer.
\newblock Parameter estimation in stochastic biochemical reactions.
\newblock {\em IEE Proceedings - Systems Biology}, 153, 2006.

\bibitem{schaefer2009pid}
Carl~F Schaefer, Kira Anthony, Shiva Krupa, Jeffrey Buchoff, Matthew Day, Timo
  Hannay, and Kenneth~H Buetow.
\newblock Pid: the pathway interaction database.
\newblock {\em Nucleic acids research}, 37(suppl 1):D674--D679, 2009.

\bibitem{Siebert06}
Heike Siebert and Alexander Bockmayr.
\newblock Incorporating time delays into the logical analysis of gene
  regulatory networks.
\newblock In {\em Computational Methods in Systems Biology}, volume 4210 of
  {\em LNCS}, pages 169--183. Springer, 2006.

\bibitem{snoussi1989qualitative}
El~Houssine Snoussi.
\newblock Qualitative dynamics of piecewise-linear differential equations: a
  discrete mapping approach.
\newblock {\em Dynamics and stability of Systems}, 4(3-4):565--583, 1989.

\bibitem{thieffry1995dynamical}
Denis Thieffry and Ren{\'e} Thomas.
\newblock Dynamical behaviour of biological regulatory networks—ii. immunity
  control in bacteriophage lambda.
\newblock {\em Bulletin of Mathematical Biology}, 57(2):277--297, 1995.

\bibitem{Thomas73}
Ren{\'e} Thomas.
\newblock Boolean formalization of genetic control circuits.
\newblock {\em Journal of Theoretical Biology}, 42(3):563 -- 585, 1973.

\bibitem{Tu2011}
C.~L. Tu, W.~Chang, and D.~D. Bikle.
\newblock {{T}he calcium-sensing receptor-dependent regulation of cell-cell
  adhesion and keratinocyte differentiation requires {R}ho and filamin {A}}.
\newblock {\em J. Invest. Dermatol.}, 131(5):1119--1128, May 2011.

\bibitem{van2013timed}
S~Van~Goethem, J-M Jacquet, Lubos Brim, and D~{\v{S}}afr{\'a}nek.
\newblock Timed modelling of gene networks with arbitrarily precise expression
  discretization.
\newblock {\em Electronic Notes in Theoretical Computer Science}, 293:67--81,
  2013.

\bibitem{yu2012hipathdb}
Namhee Yu, Jihae Seo, Kyoohyoung Rho, Yeongjun Jang, Jinah Park, Wan~Kyu Kim,
  and Sanghyuk Lee.
\newblock hipathdb: a human-integrated pathway database with facile
  visualization.
\newblock {\em Nucleic acids research}, 40(D1):D797--D802, 2012.

\end{thebibliography}



\input{partsFinal1/annexes}

%\begin{thebibliography}{4}

%\end{thebibliography}


\end{document}
